---
title: "general_workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{general_workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(duflor)
```

To quantify areas in an image, a general workflow can be accomplished by a different (set of) functions:

1. Load the image as a `cImg`-array. (`load_image()`)
2. Define the HSV-spectra to quantify. (You may either define them yourself, or modify default-spectra loaded by `spectrums <- getOption("duflor.default_hsv_spectrums")`)
3. Quantify pixels whose HSV-values fall within each spectrum defined above. (`extract_pixels_HSV()`/[`rectangularRange_HSV_iteronce_cpp()`,`rectangularRange_HSV_cpp()`,`rectangularRange_HSV()`]^[While one can certainly use the `rectangularRange_*()`-functions themselves, it is recommended to use the wrapper `extract_pixels_HSV()` to ensure input values are properly pre-processed])
4. Convert the number of pixels into an area based on a known calibration-area (`convert_pixels_to_area()`)


The alternative (and suggested) solution is to use the shiny front-end to the workflow provided above via the package [`duflor.gui`](https://github.com/Claudius-Appel/duflor.gui).
